---
title: "Intro_to_ggplot2"
author: "Whitney"
date: "12/8/2019"
output: html_document
---

## Attach packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

##Read in the data

```{r}
ca_np <- read_csv(here("R-workshop-data", "ca_np.csv"))

ci_np <- read_xlsx(here("R-workshop-data", "ci_np.xlsx"))
```

##Make a graph

1. Tell R we're using ggplot
2. Tell R what data we're using and what variables we want to plot
3. What type of graph we're creating (geom)

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```

1. Create a column graph with the data
2. Create an area graph with this data

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
geom_col()
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```

## Intro to customizing `ggplot` graphs

- color = update point or line colors
- fill = update the fill characteristics of an element
- linetype = 
- pch = value
- size
- alpha

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "coral4", linetype = "dashed")
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(color = "blue",
             pch = 17,
             size = 4, 
             alpha = 0.5)
```

I have some awesome text here, and below is my amazing finalized

```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point( color = "red",
              pch = 8,
              size = 3,
              linetype = "dashed")
```

