---
title: "Intro to readxl"
author: "Whitney"
date: "12/07/2019"
output: html_document
---

# Learning `readxl`

We are working with data in readxl and it's going to be amazing!

We are going to use packages:

- tidyverse
- readxl
- writexl
- here

```{r packages}
library(tidyverse)
# install.packages("package_name")
library(readxl)
library(writexl)
library(here)

```

### Read in fish counts

```{r}
#### Read in the fish_counts_curated.csv
fish_counts <- read_csv(here("R-workshop-data", "fish_counts_curated.csv"))

## here allows you to avoid the slashes, which can be different directions depending on whether you are using mac or windows. Also the parent directory is where the Rmd lives
```

### Read in data from a single excel worksheet

Now to read in an excel spreadsheet, we'll use the function 
`read_excel()`

read_excel("substrate_cover_curated.xlsx")

```{r}
##commeand + option + i inserts a new r code chunk

substrate_cover <- read_excel(here("R-workshop-data", "substrate_cover_curated.xlsx"),
                              skip = 4,
                              na = "-9999")
```
### Use read_excel() to read in a range of values

```{r}
carp_cover_2000 <- read_excel(here("R-workshop-data", "substrate_cover_curated.xlsx"),   range = "D12:I15",
                              col_names = c("site", "transect", "quad", "side", "type", "coverage"))
```
### How can we read in data from an Excel workbook with multiple worksheets?

```{r}
kelp_counts <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"))
```

Wait! That only read in the first worksheet.
We only want to read in the Goleta Beach sheet, named golb worksheet

```{r}
kelp_golb <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"), 
                        sheet = "golb")
```
###Read in AND combine data from multiple Excel worksheets simultaneously using purr::map_df()

```{r}
#create a path to get to the data
# control + shift + c while highlighting some code will "#" out the code, in case you want to skip it.

kelp_path <- here("R-workshop-data", "kelp_counts_curated.xlsx")

Kelp_all_sites <- kelp_path %>%
  excel_sheets()%>%
  set_names() %>%
  purr::map_df(read_excel, kelp_path)
```

### Write a file back to an Excel spreadsheet

We're going to write our "kelp_golb" object to a CSV file

```{r}
write_csv(kelp_golb, "kelp_goleta_july.csv")
# If you want to put the new csv in a new subfolder use here
#write_csv(kelp_golb, here(where you want to put it))

write_xlsx(kelp_golb, "kelp_goleta_excel.xlsx")
```

