---
title: "Pivot tables"
author: "Whitney"
date: "12/8/2019"
output: html_document
---

# Attach Libraries
```{r}

library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")
```

## Read in our data

our data is about lobsters

```{r}
lobsters <- read_xlsx(here("R-workshop-data/lobsters2.xlsx"))
```
Explore our data 

```{r}
head(lobsters)
summary(lobsters)

## skimr::skim() gives additional summary statistics and is pipable
skim(lobsters)

## ggplot of lobsters
ggplot(data = lobsters, aes(x = year))+
  geom_bar()+
  facet_wrap(~ site)
```

## Our task:

How the average size of lobsters has changed over time across sites?

Pivot tables in R are achieved by using "group_by" and "summarize"
```{r}
lobsters %>% 
  group_by(year) %>% 
  summarize(count_by_year = n())
```

What if we didn't use group_by first?

```{r}
lobsters %>% 
  summarize(count_by_something = n())

## assign our summary to a new variable 

lobsters_sum <- lobsters %>%
  summarize(count_by_something = n())
```

## group_by multiple variables

And summarize by multiple variables 
n() is a function in dplyr and gives you the number of observations in the current grouping. 

```{r}
site_year_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm =TRUE))
```

#Activity

1. Calculate the median size_mm
2. create and ggsave a plot

```{r}
site_year_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(),
            median_size_mm = median(size_mm, na.rm = TRUE))
ggplot(data = site_year_summary, aes(x = year, y = median_size_mm)) +
  geom_line(aes(color = site))

## I created a figures folder from the RStudio Files pane in order to be organized and for the code to run:
ggsave(here("figures/lobsters-line.png"))

```

#dplyr::count() for counting by group
This is a "shortcut" function

```{r}
lobsters %>%
  count(site, year)

# this is the same as 

lobsters %>%
  group_by(site, year) %>% 
  summarize(n = n()) %>% 
  ungroup()
  
```

Mutate add a new column it will make a new variable

```{r}
lobsters %>%
  mutate(size_m = size_mm/1000,
         millenia = 2000,
         observer = "Allison Horst")
  
```

Select function to subset variables by columns

lets select 
 - date
 - site
 - size_m
 
```{r}
lobsters_selected <- lobsters %>%
  mutate(size_m = size_mm/1000) %>% 
  select(date, site, size_m)
```
 
 